---
title: "HW7"
author: "Cheng Huang 2658312"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
biblio-style: asa
documentclass: article
fontsize: 11pt
bibliography: template.bib
papersize: letter
always_allow_html: yes
---

#  Normal Mixture revisite
## Find posterior density
The likelihood function given $\delta, \mu_1, \mu_2, \sigma_1, \sigma_2$ is
\[f(x|\delta, \mu_1, \mu_2, \sigma_1, \sigma_2) = \prod_{i=1}^{n}[\delta \frac{1}{\sqrt{2\pi}\sigma_1}e^{-\frac{(x_i-\mu_1)^2}{2\sigma_1^2}}+(1-\delta) \frac{1}{\sqrt{2\pi}\sigma_2}e^{-\frac{(x_i-\mu_2)^2}{2\sigma_2^2}}]\]
The prior for $\mu_1, \mu_2$ are normal $N(0, 10^2)$, so
\[\pi(\mu_1) \propto exp(-\frac{\mu_1^2}{200})\]
\[\pi(\mu_2) \propto exp(-\frac{\mu_2^2}{200})\]
The prior for $\sigma_1^2, \sigma_2^2$ are $IG(0.5, 10)$, so
\[\pi(\sigma_1^2) \propto (\sigma_1^2)^{-1.5}exp(-\frac{10}{\sigma_1^2})\]
\[\pi(\sigma_2^2) \propto (\sigma_2^2)^{-1.5}exp(-\frac{10}{\sigma_2^2})\]
Therefore the posterior distribution propotional to
\[q(\delta, \mu_1, \mu_2, \sigma_1, \sigma_2|x) \propto \prod_{i=1}^{n}[\delta \frac{1}{\sqrt{2\pi}\sigma_1}e^{-\frac{(x_i-\mu_1)^2}{2\sigma_1^2}}+(1-\delta) \frac{1}{\sqrt{2\pi}\sigma_2}e^{-\frac{(x_i-\mu_2)^2}{2\sigma_2^2}}] \times exp(-\frac{\mu_1^2}{200}) \times exp(-\frac{\mu_2^2}{200}) \]
\[\times (\sigma_1^2)^{-1.5}exp(-\frac{10}{\sigma_1^2}) \times (\sigma_2^2)^{-1.5}exp(-\frac{10}{\sigma_2^2}) \]

## MCMC
```{r 1.1, echo = TRUE, message = FALSE, warning=FALSE}

library("invgamma")
library("HI")
n <- 100
## set true value of parameters
delta <- 0.7
mu1 <- 7
mu2 <- 10
sigmasq1 <- .5
sigmasq2 <- .5

## simulation
set.seed(123)
u <- rbinom(n, prob = delta, size = 1)
d <- rnorm(n, ifelse(u == 1, mu1, mu2), ifelse(u == 1, sigmasq1, sigmasq2))

## calculate the posterior distribution
logposterior <- function(delta, mu1, mu2, sigmasq1, sigmasq2, x = d) {
  logL <- sum(log(delta * dnorm(x, mu1, sqrt(sigmasq1))+(1-delta)*dnorm(x, mu2, sqrt(sigmasq2))))
  logprior.mu1 <- dnorm(mu1, 0, 10, log = T)
  logprior.mu2 <- dnorm(mu2, 0, 10, log = T)
  logprior.sigma1 <- dinvgamma(sigmasq1, 0.5, 10, log = T)
  logprior.sigma2 <- dinvgamma(sigmasq2, 0.5, 10, log = T)
  
  return(logL + logprior.mu1 + logprior.mu2 + logprior.sigma1 + logprior.sigma2)
}

mymcmc <- function (niter, thetaInit, x = d){
  p <- length(thetaInit)
  thetaCurrent <- thetaInit
  out <- matrix(NA, niter, p)
  result <- matrix(NA, niter, p)
  for (i in 1:niter) {
      for (j in 1:p) {
        logFC <- function(thj) {
          theta <- thetaCurrent
          theta[j] <- thj
          logposterior(theta[1], theta[2], theta[3], theta[4], theta[5], d)
        }
        sp <- function(x){
          if (j==1 || j==4 || j==5) {
            ((x > 0) * (x < 1))
          } else if(j==2 || j==3 ){
            ((x > 0) * (x < 20))
          }
        }
        result[i, j] <- thetaCurrent[j] <-
        HI::arms(thetaCurrent[j], logFC, 
                 sp, 1)
      }
  }
  result
}
niter <- 2500
thetaInit <- c(0.5, 5, 5, 0.5, 0.5)
sim <- mymcmc(niter, thetaInit, d)
plot(ts(sim[,1]), main = "Delta plot")
hist(ts(sim[,1]), breaks = 50, main = "Histogram for Delta")
plot(ts(sim[,2]), main = "mu1 plot")
hist(ts(sim[,2]), breaks = 50, main = "Histogram for mu1")
plot(ts(sim[,3]), main = "mu2 plot")
hist(ts(sim[,3]), breaks = 50, main = "Histogram for mu2")
plot(ts(sim[,4]), main = "sigma1sq plot")
hist(ts(sim[,4]), breaks = 50, main = "Histogram for sigma1 square")
plot(ts(sim[,5]), main = "sigma2sq plot")
hist(ts(sim[,5]), breaks = 50, main = "Histogram for sigma2 square")
```

```{r 1.2, echo = TRUE, message = FALSE, warning=FALSE}
n <- 100
a <- 0.0; b <- 0.5
x <- rnorm(n)
y <- rpois(n, exp(a + b * x))
mydata <- data.frame(y = y, x = x)

logpost <- function(theta, data, sigma2, tau2) {
  a <- theta[1]; b <- theta[2]
  x <- data$x; y <- data$y
  return(a * sum(y) + b * sum(x * y) - exp(a) * sum(exp(b * x))
         - a^2 / 2 / sigma2 - b^2 / 2 / tau2)
}

mymcmc <- function(niter, thetaInit, data, sigma2, tau2) {
  p <- length(thetaInit)
  thetaCurrent <- thetaInit
  out <- matrix(NA, niter, p)
  for (i in 1:niter) {
    for (j in 1:p) {
      logFC <- function(thj) {
        theta <- thetaCurrent
        theta[j] <- thj
        logpost(theta, data, sigma2, tau2)
      }
      ## arms algorithm
      out[i, j] <- thetaCurrent[j] <-
        HI::arms(thetaCurrent[j], logFC, 
                 function(x) ((x > -10) * (x < 10)), 1)
    }
  }
  out
}

niter <- 500
thetaInit <- c(2, 2)
sigma2 <- tau2 <- 100
sim <- mymcmc(niter, thetaInit, mydata, sigma2, tau2)
plot(ts(sim[,1]))

plot(ts(sim[,2]))


```


```{r rej, echo = TRUE, message = FALSE, warning=FALSE}
delta <- 0.7 # true value to be estimated based on the data
n <- 100
set.seed(123)
u <- rbinom(n, prob = delta, size = 1)
x <- rnorm(n, ifelse(u == 1, 7, 10), 0.5)

mylike <- function(delta, x) {
    prod(delta * dnorm(x, 7, 0.5) + (1 - delta) * dnorm(x, 10, 0.5))
}

## simple random walk chain
myRange <- function(v, width) {
    min(1, v + width) - max(0, v - width)
}

mymcmc <- function(niter, init, x, width) {
    v <- double(niter)
    for (i in 1:niter) {
        cand <- runif(1, max(0, init - width), min(1, init + width))
        ratio <- mylike(cand, x) / myRange(cand, width) /
            mylike(init, x) * myRange(init, width)
        if (runif(1) < min(ratio, 1)) {
            v[i] <- init <- cand
        } else v[i] <- init
    }
    v
}
niter <- 2500
plot(ts(z <- mymcmc(niter, .2, x, .1)[-(1:500)]))
hist(z)
```




# Reference {-}
[jun-yan/stat-5361]https://github.com/jun-yan/stat-5361 \

